---
title: "Homework1"
author: "Allen Wang"
date: '2022-03-31'
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```
# Machine Learning Main Idea 

## Question 1
Supervised learning: Situation when for each observation of the predictor measurements $x_i, i = 1,2,...n$ there is an associated response measurement $y_i$.
Unsupervised learning: situation when for every observation $i=1,...,n$, we observe a vector of measurements $x_i$ but no associated response $y_i$.  
The difference is that whether there is a **response variable** or not.

## Question 2
Regression model is used when the response variable is quantitative; classification model is used when the response variable is qualitative. 

## Question 3 


## Question 4
Descriptive model is used to best visually emphasize a trend in data.  
Predictive model aims to predict $Y$ with minimum reducible error. It does not focus on hypothesis tests.  
Inferential model aims to test theories and causal claims, such as relationship between outcomes and predictors.

## Question 5 


## Question 6 
Predictive: the model should try to predict the binary response variable vote, which is Yes or not, based on the voter's data(predictor variables).  

Inferential: we are expected to find the relationship between the likelihood of support for the candidate and the personal contact with the candidate. A possible causal theory is tested, and the personal contact should stand out as a feature relating to voting preference.

# Exploratory Data Analysis

```{r}
library(tidyverse)
library(ggplot2)
hist(mpg$hwy,main="Histogram of highway miles per gallon",
     xlab = "Highway Miles per Gallon(hwy)")
```

```{r,eval=TRUE, echo=TRUE, include=TRUE}
ggplot(mpg, aes(x=hwy, y=cty)) + geom_point()+labs(title="hwy vs. cty")
```

```{r}
library(forcats)
mpg %>%
  ggplot(aes(x=fct_infreq(manufacturer))) + geom_bar() +
  coord_flip()
```

```{r}
ggplot(mpg, aes(x=as.character(cyl), y=hwy)) + 
  stat_boxplot(geom = "errorbar", width = 0.25) +
  geom_boxplot() + labs(x="cyl")

```

```{r}
library(corrplot)
mpg_numeric <- mpg %>% 
  select(displ,year, cyl, hwy, cty)
M = cor(mpg_numeric)
corrplot(M, method = 'number',order = 'FPC', type = 'lower', diag = FALSE)
```

